---
title: "data_insertion"
output: html_document
date: "2023-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing and loading necessary libraries

```{r}
install.packages("RMySQL")
install.packages("DBI")
#install.packages("dplyr")

library(RMySQL)
library(DBI)
#library(dplyr)
```
Connecting to DB

```{r}
con <- dbConnect(MySQL(), user = "root", password = "root", dbname = "appointment_booking", host = "localhost", bulk = TRUE, prefer_large_objects = TRUE)
dbSendQuery(con,'SET GLOBAL connect_timeout=28800')
dbSendQuery(con,'SET GLOBAL interactive_timeout=28800')
dbSendQuery(con,'SET GLOBAL wait_timeout=28800')
#dbSendQuery(con,'SET GLOBAL innodb_log_file_size=128M')
dbListTables(con)
```

Importing the raw data from CSV
```{r}
file_path <- "C:\\Users\\ameya\\Documents\\IUB\\Spring 2023\\Applied Database Technologies\\Project\\DAC_NationalDownloadableFile.csv"
df <- read.csv(file_path, header = TRUE)
filtered_df <- subset(df, st == "IN" & Med_sch != "OTHER")
nrow(filtered_df)
filtered_df <- subset(filtered_df, !is.na(Cred) & !is.na(phn_numbr) & !is.na(sec_spec_all))
nrow(filtered_df)
#head(filtered_df,5)
#sample_df <- filtered_df[sample(nrow(filtered_df), 1000), ]
#write.csv(sample_df, file = "C:\\Users\\ameya\\Documents\\IUB\\Spring 2023\\Applied Database Technologies\\Project\\sampled_data.csv", row.names = FALSE)

```

Inserting data into the different normalized tables 
Reference: https://www.geeksforgeeks.org/how-to-write-entire-dataframe-into-mysql-table-in-r/
```{r}
doctors_df <- filtered_df[, c("NPI", "frst_nm", "mid_nm", "lst_nm", "gndr")]

dbWriteTable(con, "Doctors", doctors_df[1:5], overWrite=FALSE, append=TRUE)
```

```{r}
#dbSendQuery(con,"DELETE FROM Doctors")
#dbDisconnect(con)
```

Inserting into Addresses table

```{r}
addresses_df <- sample_df[, c("adrs_id", "adr_ln_1", "adr_ln_2", "cty", "st", "zip")]
addresses_df$zip <- paste(substr(addresses_df$zip,0,5), "-",substr(addresses_df$zip,6,9))
addresses_df <- addresses_df[!duplicated(addresses_df$adrs_id), ]
#dbWriteTable(con, "Addresses", addresses_df, overWrite=FALSE, append=TRUE)
query <- paste0("INSERT INTO Addresses VALUES ", paste0("('", addresses_df$adrs_id, "', '", addresses_df$adr_ln_1, "', '", addresses_df$adr_ln_2, "', '", addresses_df$cty, "', '", addresses_df$st, "', '", addresses_df$zip, "')", collapse = ", "))
dbExecute(con, query)
```
Inserting into Contacts table

```{r}
contacts_df <- sample_df[, c("NPI", "phn_numbr")]
contacts_df$phn_numbr <- paste("(", substr(contacts_df$phn_numbr,0,3), ") -", substr(contacts_df$phn_numbr,4,6), "-",  substr(contacts_df$phn_numbr,7,10))
head(contacts_df,5)
query <- paste0("INSERT INTO Contacts VALUES ", paste0("(", contacts_df$NPI, ", '", contacts_df$phn_numbr, "')", collapse = ", "))
dbExecute(con, query)
```

Inserting into Services table

```{r}
services_df <- sample_df[, c("NPI", "Telehlth", "ind_assgn", "grp_assgn")]
services_df <- services_df[!duplicated(services_df$NPI), ]

query <- paste0("INSERT INTO Services VALUES ", paste0("(", services_df$NPI, ", '", services_df$Telehlth, "', '", services_df$ind_assgn, "', '", services_df$grp_assgn, "')", collapse = ", "))
dbExecute(con, query)
```

Inserting into Schools table

```{r}
schools <- sample_df$Med_sch
schools <- unique(schools)

query <- paste0("INSERT INTO Schools (SchoolName) VALUES ", paste0("('", schools, "')", collapse = ", "))
dbExecute(con, query)
```

Inserting into Education table

```{r}
education_df <- sample_df[, c("NPI", "Med_sch", "Cred", "Grd_yr")]
dbWriteTable(con, "Education_temp", education_df, APPEND = TRUE)

```

Inserting into DoctorClinics table

```{r}
clinics_df <- sample_df[, c("NPI", "adrs_id")]
clinics_df <- clinics_df[!duplicated(clinics_df$NPI), ]

query <- paste0("INSERT INTO DoctorClinics VALUES ", paste0("(", clinics_df$NPI, ", '", clinics_df$adrs_id, "')", collapse = ", "))
dbExecute(con, query)
```

Inserting into DoctorOrganizations table

```{r}
doctor_organizations_df <- sample_df[, c("NPI", "org_pac_id", "adrs_id")]
doctor_organizations_df <- doctor_organizations_df[complete.cases(doctor_organizations_df$org_pac_id), ]
duplicates_df <- doctor_organizations_df[!duplicated(doctor_organizations_df$adrs_id), ]
duplicates_df$IsHospital <- rep(1, nrow(duplicates_df))
uniques_df <- doctor_organizations_df[duplicated(doctor_organizations_df$adrs_id), ]

query <- paste0("INSERT INTO DoctorOrganizations VALUES ", paste0("(", duplicates_df$NPI, ", '", duplicates_df$org_pac_id, "', '", duplicates_df$IsHospital, "')", collapse = ", "))
dbExecute(con, query)
```

Inserting into Specialties table

```{r}
specialties <- c(sample_df$pri_spec, sample_df$sec_spec_1, sample_df$sec_spec_2, sample_df$sec_spec_3, sample_df$sec_spec_4)
specialties <- unique(specialties)

query <- paste0("INSERT INTO Specialties (SpecialtyName) VALUES ", paste0("('", specialties, "')", collapse = ", "))
dbExecute(con, query)
```

Inserting into DoctorSpecialties table

```{r}
doctor_specialties_df <- sample_df[, c("NPI", "pri_spec", "sec_spec_1", "sec_spec_2", "sec_spec_3", "sec_spec_4")]

primary_specialty_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$pri_spec)
primary_specialty_df$IsPrimary <- rep(1, nrow(primary_specialty_df))


secondary_specialty1_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_1)
secondary_specialty1_df[complete.cases(secondary_specialty1_df$specialty), ]
secondary_specialty1_df$IsPrimary <- rep(0, nrow(secondary_specialty1_df))

secondary_specialty2_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_2)
secondary_specialty2_df <- secondary_specialty2_df[complete.cases(secondary_specialty2_df$specialty), ]
secondary_specialty2_df$IsPrimary <- rep(0, nrow(secondary_specialty2_df))

secondary_specialty3_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_3)
secondary_specialty3_df <- secondary_specialty3_df[complete.cases(secondary_specialty3_df$specialty), ]
secondary_specialty3_df$IsPrimary <- rep(0, nrow(secondary_specialty3_df))

secondary_specialty4_df <- data.frame(NPI = doctor_specialties_df$NPI, specialty = doctor_specialties_df$sec_spec_4)
secondary_specialty4_df <- secondary_specialty4_df[complete.cases(secondary_specialty4_df$specialty), ]
secondary_specialty4_df$IsPrimary <- rep(0, nrow(secondary_specialty4_df))

nrow(primary_specialty_df)
combined_df <- rbind(primary_specialty_df, secondary_specialty1_df, secondary_specialty2_df, secondary_specialty3_df, secondary_specialty4_df)
nrow(combined_df)
dbWriteTable(con, "DoctorSpecialties_temp", combined_df, APPEND = TRUE)
```